Reading main files and preparing data sets

### 2025-06-01 data-raw/03-preprocessing-main-files.Rmd
Updating package documentation and 2022 TUS GSS variables.

```{r reading-2022-main-file}
# Main File
gss_m_2022 <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_main_2022_pumf_v1.sav"), to.data.frame = TRUE) # Original file
```

@@

```{r filtering-variables-2022}
gss_main_2022 <- gss_m_2022 %>%
  dplyr::select("PUMFID", "WGHT_PER","SURVMNTH", "PRV", "LUC_RST", "REGION",  "HSDSIZEC","AGEGR10", "GENDER2",
"MARSTAT", "PHSDFLG", "AGEPRC",  "GENPR2",  "COUPGEN2","CHH0017C","CHH0004C","CHH0514C","CHH1517C","CXRFLAG",
"PARNUM",  "SENFLAG", "MULTIGEN","DDEV_FL", "DDEX_FL", "DFLEX_FL","DHEAR_FL","DLRN_FL", "DMEM_FL", "DMENT_FL",
 "DMOB_FL", "DPAIN_FL","DUNK_FL", "DVIS_FL", "DDIS_FL", "DTYPER",  "DCLASS",  "TCS_110", "TCS_120", "TCS_130",
 "TCS_140", "TCS_150", "TCS_160", "TCS_170", "TCS_180", "TCS_190", "TCS_200",  "GEN_01",  "GEN_02",  "SRS_10",  "MSS_130", "ACT7DAYC","MRW_05C", "MRW_D40A",
"MRW_D40B","WET_120",  "WHW_230", "WHWD140G","TLWK_01A","TLWK_06","WLYD170G",
"ATT_120", "ATT_130", "ATT_131",
"ATT_150C","ATT_160", "ATT_170",
"CTW_140A","CTW_140B","CTW_140C","CTW_140D",
"CTW_140E","CTW140GR","CTW_140I","CTW_150G","CTW_160", "CTW_170", "CTW_180", "CTW_190", "CTW_210", "CTW_215",
"ED_05",  "MAP_110C","MAP_D40A","DOR_110C","IM_01A",  "YRARRI_C","BPP_01A", "BPR_20", "ABM_01A",
"AIP_01A", "VISMINFL","VISPRFL", "REL_02",  "RELIGFLG","INC_C",  
"FAMINC_C","GTU_110", "GTU_130",  "DDAY",  "DVTDAY") %>% 
  mutate(PUMFID = as.numeric(as.character(PUMFID)),
         WGHT_PER = as.numeric(as.character(WGHT_PER)))
```

>>>>

```{r reading-2022-main-file}
# Main File
gss_m_2022 <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_main_2022_pumf_v1.sav"), to.data.frame = TRUE) # Original file

gss_m_2022_b <- gss_m_2022
gss_m_2022 <- gss_m_2022_b
```

@@

```{r filtering-variables-2022}
gss_main_2022 <- gss_m_2022 %>%
  dplyr::select("PUMFID", "WGHT_PER","SURVMNTH", "PRV", "LUC_RST", "REGION",  "HSDSIZEC","AGEGR10", "GENDER2",
"MARSTAT", "PHSDFLG", "AGEPRC",  "GENPR2",  "COUPGEN2","CHH0017C","CHH0004C","CHH0514C","CHH1517C","CXRFLAG",
"PARNUM",  "SENFLAG", "MULTIGEN","DDEV_FL", "DDEX_FL", "DFLEX_FL","DHEAR_FL","DLRN_FL", "DMEM_FL", "DMENT_FL",
 "DMOB_FL", "DPAIN_FL","DUNK_FL", "DVIS_FL", "DDIS_FL", "DTYPER",  "DCLASS",  "TCS_110", "TCS_120", "TCS_130",
 "TCS_140", "TCS_150", "TCS_160", "TCS_170", "TCS_180", "TCS_190", "TCS_200",  "TIMECR", "GEN_01",  "GEN_02",  "SRS_10",  "MSS_130", "ACT7DAYC","MRW_05C", "MRW_D40A",
"MRW_D40B","WET_120",  "WHW_230", "WHWD140G","TLWK_01A","TLWK_06","WLYD170G",
"ATT_120", "ATT_130", "ATT_131",
"ATT_150C","ATT_160", "ATT_170",
"CTW_140A","CTW_140B","CTW_140C","CTW_140D",
"CTW_140E","CTW140GR","CTW_140I","CTW_150G","CTW_160", "CTW_170", "CTW_180", "CTW_190", "CTW_210", "CTW_215",
"ED_05",  "MAP_110C","MAP_D40A","DOR_110C","IM_01A",  "YRARRI_C","BPP_01A", "BPR_20", "ABM_01A",
"AIP_01A", "VISMINFL","VISPRFL", "REL_02",  "RELIGFLG","INC_C",  
"FAMINC_C","GTU_110", "GTU_130",  "DDAY",  "DVTDAY") %>% 
  mutate(PUMFID = as.numeric(as.character(PUMFID)),
         WGHT_PER = as.numeric(as.character(WGHT_PER)))
```

```{r LUC_RST-2022}
gss_main_2022 <- gss_main_2022 %>% 
  mutate( # LUC_RST:  Population centre indicator
    LUC_RST = 
      case_when(LUC_RST == "Larger urban population centres (CMA/CA)" ~ "CMA/CA",
                     LUC_RST == "Rural areas/small population centres (non-CMA/CA)" ~ "non CMA/CA",
                     LUC_RST == "Prince Edward Island" ~ "Prince Edward Island"),
    LUC_RST = factor(LUC_RST))

summary(gss_main_2022$LUC_RST)
```
```{r HSDSIZEC-2022}
values_ordered <- c("One person", 
           "Two persons",
           "Three persons",
           "Four persons",
           "Five persons or more")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# HSDSIZEC:  Household size of respondent (capped)
    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$HSDSIZEC)
```
```{r AGEGR10-2022}
values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# AGEGR10:  Age group of respondent (groups of 10)
    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE), 
    AGEPRC = factor(AGEPRC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$AGEGR10)
```
```{r AGEPRC-2022}
values_ordered <- c("15 to 34 years","35 to 44 years",
                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# AGEPRC:  Age group of respondent’s spouse/partner (collapsed)
    AGEPRC = factor(AGEPRC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$AGEPRC)
```

```{r CHH0017C-2022}
summary(gss_main_2022$CHH0017C)

values_ordered <- c("No children","One child",
                    "Two children", "Three children or more")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# CHH0017C:  Number of child(ren) in household - 0 to 17 years (capped)
    CHH0017C = factor(CHH0017C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$CHH0017C)
```
```{r CHH0004C-2022}
summary(gss_main_2022$CHH0004C)

values_ordered <- c("No children","One or more children")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# CHH0004C:  Number of child(ren) in household - 0 to 4 years (capped)
    CHH0004C = factor(CHH0004C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$CHH0004C)
```

```{r CHH0514C-2022}
summary(gss_main_2022$CHH0514C)

values_ordered <- c("No children","One or more children")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# CHH0514C:  Child(ren) in household - 5 to 14 years (flag)
    CHH0514C = factor(CHH0514C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$CHH0514C)
```

```{r CHH1517C-2022}
summary(gss_main_2022$CHH1517C)

values_ordered <- c("No children","One or more children")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# CHH1517C:  Child(ren) in household - 15 to 17 years (flag)
    CHH1517C = factor(CHH1517C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$CHH1517C)
```

```{r PARNUM-2022}
summary(gss_main_2022$PARNUM)

values_ordered <- c("None","One parent","Two parents")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# PARNUM: Number of parents the respondent has in household
    PARNUM = factor(PARNUM, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$PARNUM)
```
```{r DTYPER-2022}
summary(gss_main_2022$DTYPER)

values_ordered <- c("Does not have a disability",
                    "Has one disability type",
                    "Has two or three disability types",
                    "Has more than three disability types")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# DTYPER: Disability type counter - Grouped
    DTYPER = factor(DTYPER, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$DTYPER)
```

```{r DCLASS-2022}
summary(gss_main_2022$DCLASS)

values_ordered <- c("Does not have a disability",
                    "Unknown severity",
                    "Mild",
                    "Moderate",
                    "Severe",
                    "Very severe")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# DCLASS: Global (disability) severity class
    DCLASS = factor(DCLASS, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$DCLASS)
```

```{r TIMECR-2022}
summary(gss_main_2022$TIMECR)

values_ordered <- c("Zero Yes codes",
                    "One Yes code",
                    "Two Yes codes",
                    "Three Yes codes",
                    "Four Yes codes",
                    "Five Yes codes",
                    "Six Yes codes",
                    "Seven Yes codes",
                    "Eight Yes codes",
                    "Nine Yes codes",
                    "Ten Yes codes")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# TIMECR: Time crunch indicator
    TIMECR = factor(TIMECR, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$TIMECR)
```

```{r GEN_01-2022}
summary(gss_main_2022$GEN_01)

values_ordered <- c("Poor",
                    "Fair",
                    "Good",
                    "Very good",
                    "Excellent")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# GEN_01: Self rated health - In general, how is your health?
    GEN_01 = factor(GEN_01, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$GEN_01)
```

```{r GEN_02-2022}
summary(gss_main_2022$GEN_02)

values_ordered <- c("Poor",
                    "Fair",
                    "Good",
                    "Very good",
                    "Excellent")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# GEN_02: Self rated mental health - In general, how is your mental health?
    GEN_02 = factor(GEN_02, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$GEN_02)
```

```{r SRS_10-2022}
summary(gss_main_2022$SRS_10)

values_ordered <- c("Not at all stressful",
                    "Not very stressful",
                    "A bit stressful",
                    "Quite a bit stressful",
                    "Extremely stressful")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# SRS_10: Self Rated Stress - Amount - Thinking about the amount of stress in your life, how would you describe most of your days?
    SRS_10 = factor(SRS_10, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$SRS_10)
```

```{r TLWK_06-2022}
summary(gss_main_2022$TLWK_06)

values_ordered <- c("1 - Not important at all",
                    "2",
                    "3",
                    "4",
                    "5 - Very important")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# TLWK_06: Telework aspects - Avoid the commute. How important is the opportunity not to commute every day to your workplace?
    TLWK_06 = factor(TLWK_06, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$TLWK_06)
```

```{r WLYD170G-2022}
summary(gss_main_2022$WLYD170G)

values_ordered <- c("Respondent works from home",
                    "0 to 4 kilometers",
                    "5 to 9 kilometers",
                    "10 to 14 kilometers",
                    "15 to 19 kilometers",
                    "20 to 24 kilometers",
                    "25 to 29 kilometers",
                    "30 to 39 kilometers",
                    "40 to 49 kilometers",
                    "50 kilometers or more")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# WLYD170G: Distance from residence to workplace (grouped)
    WLYD170G = factor(WLYD170G, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$WLYD170G)
```

```{r ATT_120-2022}
summary(gss_main_2022$ATT_120)

values_ordered <- c("All the time",
                    "Some of the time",
                    "Rarely",
                    "Never")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# ATT_120: Vehicle at respondent’s disposal. How often do you have a vehicle at your disposal?
    ATT_120 = factor(ATT_120, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$ATT_120)
```

```{r ATT_130-2022}
summary(gss_main_2022$ATT_130)

values_ordered <- c("All the time",
                    "Some of the time",
                    "Rarely",
                    "Never")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# ATT_130: Need someone else to drive - Frequency. How often do you need someone to drive you somewhere for an appointment, errand or something else?
    ATT_130 = factor(ATT_130, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$ATT_130)
```

```{r ATT_131-2022}
summary(gss_main_2022$ATT_131)

values_ordered <- c("All the time",
                    "Some of the time",
                    "Rarely",
                    "Never")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# ATT_131: Someone else can drive when needed - Frequency. How often can someone drive you when you need it?
    ATT_131 = factor(ATT_131, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$ATT_131)
```
```{r ATT_170-2022}
summary(gss_main_2022$ATT_170)

values_ordered <- c("Not in past 12 months",
                    "A few times a year",
                    "A few times a month",
                    "Several times a week",
                    "Every day")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# ATT_170: Use of public transportation - Past 12 months.
    ATT_170 = factor(ATT_170, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$ATT_170)
```

```{r ED_05-2022}
summary(gss_main_2022$ED_05)

values_ordered <- c("Less than high school diploma or its equivalent",
                    "High school diploma or a high school equivalency certificate",
                    "Trades certificate or diploma",
                    "College, CEGEP or other non-university certificate or diplom",
                    "University certificate or diploma below the bachelor's level",
                    "Bachelor's degree",
                    "University certificate, diploma, or degree above the bachelo")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# ED_05: Educational attainment - Highest degree (7 categories)
    ED_05 = factor(ED_05, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$ED_05)
```

```{r REL_02-2022}
summary(gss_main_2022$REL_02)

values_ordered <- c("Not at all",
                    "Once or twice a year",
                    "At least three times a year",
                    "At least once a month",
                    "At least once a week")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# REL_02: Frequency of religious participation - Past 12 months.
    REL_02 = factor(REL_02, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$REL_02)
```
```{r INC_C-2022}
summary(gss_main_2022$INC_C)

values_ordered <- c("Less than $25,000",
                    "$25,000 to $49,999",
                    "$50,000 to $74,999",
                    "$75,000 to $99,999",
                    "$100,000 and over")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# INC_C: Income - Total (before tax) (grouped).
    INC_C = factor(INC_C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$INC_C)
```

```{r FAMINC_C-2022}
summary(gss_main_2022$FAMINC_C)

values_ordered <- c("Less than $50,000",
                    "$50,000 to $74,999",
                    "$75,000 to $99,999",
                    "$100,000 to $124,999",
                    "$125,000 and over")

gss_main_2022 <- gss_main_2022 %>% 
  mutate(# FAMINC_C: Family Income - Total (before tax) (grouped).
    FAMINC_C = factor(FAMINC_C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2022$FAMINC_C)
```
```{r GTU_110-2022}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never")

gss_main_2022 <- gss_main_2022 %>% 
  mutate( # GTU_110: How often do you feel rushed? Would you say it is...?
    GTU_110 = factor(GTU_110, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2022$GTU_110)
```

```{r GTU_130-2022}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never")

gss_main_2022 <- gss_main_2022 %>% 
  mutate( # GTU_130: How often do you feel you have time on your hands that you don’t know what to do with?
    GTU_130 = factor(GTU_130, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2022$GTU_130)
```
